Results

### 2024-08-16 analysis/results.Rmd
More upsets :-)

```{r r_setup}
library(bigrquery)
library(tidyverse)
library(hoaddata)
#' Connect to DB 
#' 
#' BigQuery HOAD (Crossref, OpenAlex and open friends)
bq_con <- dbConnect(
  bigrquery::bigquery(),
  billing = "subugoe-collaborative",
  project = "subugoe-collaborative"
)
```

@@

```{r upset_data, fig.asp=0.8}
cr_active_upset <- cr_df |>
  filter(issn_l %in% active_jns$issn_l) |>
  group_by(issn_l) |>
  summarise(n_cr = n_distinct(doi)) |>
  mutate(Crossref = TRUE)
wos_active_upset <- wos_active_core |>
  group_by(issn_l) |>
  summarise(n_wos = n_distinct(item_id)) |>
  mutate(WoS = TRUE)
scp_active_upset <- scp_active_core |>
  group_by(issn_l) |>
  summarise(n_scp = n_distinct(item_id)) |>
  mutate(Scopus = TRUE)

@@

library(ComplexUpset)
jn_upset <- left_join(jn_upset_, esac, by = "issn_l")

@@

>>>>

```{r r_setup}
library(bigrquery)
library(tidyverse)
library(hoaddata)
library(patchwork)

#' Connect to DB 
#' 
#' BigQuery HOAD (Crossref, OpenAlex and open friends)
bq_con <- dbConnect(
  bigrquery::bigquery(),
  billing = "subugoe-collaborative",
  project = "subugoe-collaborative"
)
```

@@

```{r upset_data, fig.asp=0.8}
cr_active_upset <- cr_df |>
  filter(issn_l %in% active_jns$issn_l) |>
  group_by(issn_l) |>
  summarise(n_cr = n_distinct(doi)) |>
  mutate(Crossref = TRUE)
wos_active_upset <- wos_active_jns |>
  group_by(issn_l) |>
  summarise(n_wos = n_distinct(item_id)) |>
  mutate(WoS = TRUE)
scp_active_upset <- scp_active_jns |>
  group_by(issn_l) |>
  summarise(n_scp = n_distinct(item_id)) |>
  mutate(Scopus = TRUE)

@@

library(ComplexUpset)
jn_upset <- inner_join(jn_upset_, esac, by = "issn_l") |>
  distinct(issn_l, .keep_all = TRUE)

@@

### DOI coverage

```{r doi_coverage_check}
jn_overlap <- jn_upset |> 
  filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE)

miss_doi <- cr_df |>
  filter(issn_l %in% jn_overlap$issn_l) |> 
  filter(!doi %in% wos_active_core$doi)

# Check if it is in WOS (full)
wos_type_diff <- wos_jct_items_df |>
  filter(doi %in% miss_doi$doi)

miss_doi |>
  filter(!doi %in% wos_type_diff$doi) |>
  sample_n(10)
```

### Add Article Coverage

- hoaddata: Crossref exclunding paratext and proceedings as indicated by pattern macthing. Patterns mostly from Unpaywall / OpenAlex
- WoS: All DOIs / DOIS only article and reviews
- Scopus: All DOIs / DOIS only article and reviews

```{r fig.asp=0.8}
 doi_overlap_df <- cr_df |>
  filter(issn_l %in% active_jns$issn_l) |>
  mutate(in_wos_and_scopus = ifelse(doi %in% wos_active_jns$doi & doi %in% scp_active_jns$doi, 1, 0),
         in_scopus = ifelse(doi %in% scp_active_jns$doi, 1, 0),
         in_wos =  ifelse(doi %in% wos_active_jns$doi, 1, 0))
doi_overlap <- doi_overlap_df |>
  group_by(issn_l) |>
  summarise(
    doi_in_wos_and_scopus = sum(in_wos_and_scopus),
    doi_in_scopus = sum(in_scopus),
    doi_in_wos = sum(in_wos))

jn_upset_articles <- jn_upset |> 
  left_join(doi_overlap, by = "issn_l")

### Plot
size = get_size_mode('exclusive_intersection')

upset_overview <- ComplexUpset::upset(
  jn_upset_articles,
  c("Crossref", "WoS", "Scopus"),
  # Only journals in Crossref
  min_size = 30,
  width_ratio = 0.4,
  set_sizes=FALSE,
  stripes = "transparent",
  wrap = TRUE,
  name = "Data Source Coverage",
  base_annotations = list(
    "Journals" = (
      ComplexUpset::intersection_size(
        counts = TRUE,
        text = list(size = 9 / .pt),
        text_mapping = aes(
          label = paste0(!!upset_text_percentage(), '\n(', !!size, ')'),
          colour = ifelse(!!size > 5000, 'on_bar', 'on_background'),
          y = ifelse(!!size > 5000, !!size - 4000, !!size)
        )
      ) +
        scale_y_continuous(labels = scales::number_format(big.mark = ","))
    )
  ),
  annotations = list("Journal size\n(log scale)" = (
    ggplot(mapping = aes(x = intersection, y = n)) +
      geom_boxplot(outliers = FALSE) +
      scale_y_log10(labels =  scales::number_format(big.mark = ","))),
    "Articles" = (ggplot(mapping = aes(y = n_cr)) +
                        geom_bar(stat = "identity", color = "grey80", aes(fill = "in Crossref only")) +
                        geom_bar(aes(y = doi_in_wos_and_scopus, fill = "Shared"), stat = "identity", color = "#0093c7") +
                        geom_bar(aes(y = doi_in_wos, fill = "Shared"), stat = "identity", color = "#0093c7") +
                        geom_bar(aes(y = doi_in_scopus, fill = "Shared"), stat = "identity", color = "#0093c7") +
                        scale_fill_manual(values = c(`in Crossref only` = "grey80", `Shared` = "#0093c7"), name = "") +
                        scale_y_continuous(labels =  scales::number_format(big.mark = ",")) +
                        theme(
    legend.position = "top",
    legend.justification = "right"
  ) +
    guides(fill = "none")
    )),
   matrix = (
        intersection_matrix(
            geom = geom_point(
                shape = "square",
                size = 2
            )))
        ) 
upset_overview
```

```{r dois_over_the_years}
doi_coverage_over_years <- doi_overlap_df |>
  group_by(cr_year) |>
  summarise(dois = n_distinct(doi),
            in_wos_and_scopus = sum(in_wos_and_scopus, na.rm = TRUE),
            in_scopus = sum(in_scopus, na.rm = TRUE),
            in_wos = sum(in_wos)) |>
  pivot_longer(cols = c(dois, in_wos_and_scopus, in_scopus, in_wos)) |>
  mutate(cr_year = gsub("^20", "'", as.character(cr_year)))

# dois all
dois_all_by_year <- doi_coverage_over_years |>
  filter(name == "dois") |>
  select(-name)

coverage_over_year_plot <- doi_coverage_over_years |>
  filter(name != "dois") |>
  # Fix order of panel
  mutate(name = factor(name, levels = c("in_wos_and_scopus", "in_scopus", "in_wos"))) |>
  ggplot(aes(x = cr_year, y = value)) +
  geom_bar(data = dois_all_by_year, aes(fill = "in Crossref only"), color = "transparent", stat = "identity") +
  geom_bar(aes(fill = "Shared"), color = "transparent", stat = "identity") +
  facet_wrap( ~ name, ncol = 1) +
  scale_fill_manual(values = c("grey80", "#0093c7"), name = "") +
  scale_y_continuous(labels =  scales::number_format(big.mark = ",")) +
  labs(x = "Year published", y = "Articles with DOI") + 
  theme_minimal() +
  theme(legend.position = "top",
        legend.justification = "right") 
coverage_over_year_plot
```

```{r coverage_upset_patch, fig.asp=0.85}
design = "AB"
wrap_plots(
  A = upset_overview,
  B = coverage_over_year_plot,
  design = design,
  widths = c(3.5,1)
) & plot_annotation(tag_levels = 'A') 
```
